Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-09-20 SDSS Journal Paper/Paper Draft.Rmd
one last \db comment

abstract: >
  Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
keywords:
# alphabetical; excluding anything in the title already; lower case
  - Interactive Dashboards
  - Exploratory Data Analysis (EDA)
  - Parallel Coordinate Plots (PCP)
  - Guided Discovery Learning (GDL)
bibliography: jdsart-sample.bib
output: 
  jds.rmd::pdf_article:
    keep_tex: true
---

>>>>

abstract: >
  Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will use Iowa data to  explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
keywords:
# alphabetical; excluding anything in the title already; lower case
  - Interactive Dashboards
  - Exploratory Data Analysis (EDA)
  - Parallel Coordinate Plots (PCP)
  - Guided Discovery Learning (GDL)
bibliography: jdsart-sample.bib
output: 
  jds.rmd::pdf_article:
    keep_tex: true
---

### 2022-09-20 SDSS Journal Paper/Paper-Draft.tex
one last \db comment

\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}

@@

\begin{abstract}
Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
\end{abstract}

@@

{\textcolor{green}{\sout{Abstract}}}

{\textcolor{green}{\sout{Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst. }}}

@@

Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns. The {\textcolor{green}{\sout{Index of Relative Rurality (}}}IRR{\textcolor{green}{\sout{)}}} is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework {\textcolor{green}{that}} leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours \citep{tourr}, require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s \citep{dOcagne:1885}, parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show ``blocks'' of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables \citep{Bendix:2005}; modifications of this design, such as common-angle plots \citep{Hofmann:2013}, address the issues which arise due to line-width illusions \citep[\citet{sineillusion}]{sine}. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}. In this paper, we use the \texttt{ggpcp} package {\textcolor{green}{in R [@Rlanguage]}}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's {\textcolor{green}{Quality of Life (QoL)}} and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

\begin{itemize}
\tightlist
\item
  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with {\textcolor{green}{QoL, grouped into four domains: physical, psychological, social, and environmental}}. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by {\textcolor{green}{\sout{quality of life} QoL}} factor \citep{petersCommunityResiliencyDeclining2019}.
\item
  Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light \(\alpha\) grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-\(\alpha\) value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
\item
  We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
\end{itemize}

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the {\textcolor{green}{applet}} and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining {\textcolor{green}{\sout{quality of life} QoL}} amid shrinking populations.

@@

EDA can be {\textcolor{green}{\sout{a}}} difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.

>>>>

\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{blue}{#1}}}

@@

\begin{abstract}
Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will use Iowa data to explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
\end{abstract}

@@

@@

Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns. The IRR is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework that leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours \citep{tourr}, require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s \citep{dOcagne:1885}, parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show ``blocks'' of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables \citep{Bendix:2005}; modifications of this design, such as common-angle plots \citep{Hofmann:2013}, address the issues which arise due to line-width illusions \citep[\citet{sineillusion}]{sine}. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}. In this paper, we use the \texttt{ggpcp} package in R \citep[\citet{ggpcp}]{Rlanguage}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's Quality of Life (QoL) and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

\begin{itemize}
\tightlist
\item
  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with QoL, grouped into four domains: physical, psychological, social, and environmental. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by QoL factor \citep{petersCommunityResiliencyDeclining2019}.
\item
  Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light \(\alpha\) grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-\(\alpha\) value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
\item
  We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
\end{itemize}

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the applet and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining QoL amid shrinking populations.

@@

EDA can be difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.
